---
title: "Random Forest"
author: "Anton"
output: 
        html_document:
          toc: true
          theme: united
          toc_float: true
          toc_depth: 5
          highlight: tango
          
---
<head>
			<meta charset="utf-8"/>
			<title>Case Study Non-linear Models</title>
			<style> 
      	body {
      		padding-left: 2.5%;
      		padding-right: 10%;
      	}

				h1 {
					font-size: 32px;
				}

				h2 {
					font-size: 28px;
					color: darkblue;
				}
				h3 {
					font-size: 22px;
					color: darkblue;
				}
				h4 {
				  font-size: 18px;
				}
				h5 {
				  font-size: 16px;
				}
			</style>
		</head>
<body>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("/Users/antonehret/Desktop/datensets"))
```
<h1>Random Forest in R</h2>
<br/>
<h2>Preparations</h2>
<br/>
<h3>Libraries</h3>
<br/>
```{r message=FALSE, tidy=TRUE}
#' *Libraries*
    library(randomForest)
    library(caret)
    library(rpart)
```
<br/>
<h3>Load Data</h3>
```{r message=FALSE, tidy=TRUE}
#' *Load Data*
    train=read.csv("train.csv")
    test=read.csv("test.csv")
#' Add "Survived" column to test, to help combine with train data
    test$Survived=NA
#' Combine train and test
    combi=rbind(train,test)
```
<br/>
<h3>Data Wrangling</h3>
```{r message=FALSE, tidy=TRUE}
#' *Data Wrangling*
#' Handling @Name
#' Convert names to character
    combi$Name<-as.character(combi$Name)
#' test of how strsplit works
    strsplit(combi$Name[1],split='[,.]')
    strsplit(combi$Name[1],split='[,.]')[[1]]
    strsplit(combi$Name[1],split='[,.]')[[1]][2]
    
#' apply function to dataset. This will isolate title for all rows
    combi$Title <- sapply(combi$Name, FUN=function(x){strsplit(x,split='[,.]')[[1]][2]})
    
#' Handling @title
#' Apply function to dataset, this will isolate title for all rows
    combi$Title <- sapply(combi$Name, FUN=function(x){strsplit(x,split='[,.]')[[1]][2]})
#' remove empty spaces from the 'Title' field
    combi$Title=gsub(' ','',combi$Title)
#' Review contents of 'Title' field
    table(combi$Title)
#' Reduce 'Title' contenst into fewer categories
    combi$Title[combi$Title %in% c('Mme', 'Mlle')] <- 'Mlle'
    combi$Title[combi$Title %in% c('Capt', 'Don', 'Major', 'Sir')] <- 'Sir'
    combi$Title[combi$Title %in% c('Dona', 'Lady', 'theCountess', 'Jonkheer')] <- 'Lady'
#' Change Title to a factor
    combi$Title <- factor(combi$Title)

#' Handling @sibling and @parent/child
    combi$FamilySize <- combi$SibSp + combi$Parch + 1
    
#' Identifying @families by combining last name and family size
#' identify surname
    combi$Surname <- sapply(combi$Name, FUN=function(x) {strsplit(x, split='[,.]')[[1]][1]})
#' combine with family size
    combi$FamilyID <- paste(as.character(combi$FamilySize), combi$Surname, sep="")
#' Categorize family size less than 2 as small
    combi$FamilyID[combi$FamilySize <= 2] <- 'Small'
#' Review results
    table(combi$FamilyID)
#' Further consolidate results (some families may have different last names)
    famIDs <- data.frame(table(combi$FamilyID))
    famIDs <- famIDs[famIDs$Freq <= 2,]
    combi$FamilyID[combi$FamilyID %in% famIDs$Var1] <- 'Small'
    combi$FamilyID <- factor(combi$FamilyID)
```
<br/>
<h3Taking care of missing values</h3>
```{r}
    library(DataExplorer)
    plot_missing(combi)
```
<br/>
The "Age" variable has a few missing values.
To use the randomForest package in R, there should be no missing values.
A quick way of dealing with missing values is to replace them with either the mean or the median of the non-missing values for the variable. In this example, we are replacing the missing values with a prediction, using decision trees.
<br/>
```{r, tidy=TRUE}
    library(rpart)
    Agefit <- rpart(Age ~ Pclass + Sex + SibSp + Parch + Fare + Embarked + Title + FamilySize
                    , data=combi[!is.na(combi$Age),]
                    , method="anova")
    combi$Age[is.na(combi$Age)] <- predict(Agefit, combi[is.na(combi$Age),])
    plot_missing(combi) #the missing values from $survived are due to the training dataset
```
<br/>
Check for other missing variables
```{r , tidy=TRUE}
#' @Embarked has two blank variables
#' They are identified using the "which" command
    which(combi$Embarked == '')
#' rows 62 and 830 have the blank values for Embarked
#' they are replaced with the mode of all the values for Embarked, which is 'S'
    combi$Embarked[c(62,830)] = "S"
#' convert 'Embarked' to a factor
    combi$Embarked <- factor(combi$Embarked)
    
#' @Fare has one NA value
    which(is.na(combi$Fare))
#Replace with median
    combi$Fare[1044] <- median(combi$Fare, na.rm=TRUE)
```

<h3>Factor levels for Random Forests</h3>
<br/>
Random Forests in R can only digest factors with up to 32 levels.
If any factor variable has more than 32 levels, the levels need to be redefined to be <=32 or the variable needs to be converted into a continuous one.
Here we will redefine the levels:
<br/>
```{r , tidy=TRUE}
#' *Checking the Nr of levels*
    str(combi)
#' We see that the FamilyID has too many levels (62)
```

```{r}
#' *Taking care of FamilyID* 
#' increase the definition of Small from 2 to 3
    combi$FamilyID2 <- combi$FamilyID
    combi$FamilyID2 <- as.character(combi$FamilyID2)
    combi$FamilyID2[combi$FamilySize <= 3] <- 'Small'
    combi$FamilyID2 <- factor(combi$FamilyID2)
    str(combi$FamilyID2) #We now have 22 levels
```
<br/>
<h3>Data splitting</h3>
Splitting this new dataset back into train and test datasets
```{r , tidy=TRUE}
    train <- combi[1:891,]
    test <- combi[892:1309,]
```
<br/>
<br/>
<h2>Modelling</h2>
```{r, tidy=TRUE}
    set.seed(123)
    fit <- randomForest(as.factor(Survived) ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked + Title + FamilySize + FamilyID2
                        , data = train
                        , importance = TRUE #allows us to inspect variable importance
                        , ntree = 2000) #enables specifying how many trees we want to grow
    summary(fit)
    varImpPlot(fit)
    summary(combi$Title)
```

    MeanDecreaseAccuracy : Tells us how much the accuracy decreases without the variable on the Y-axis.
    'Title' causes the most decrease and is therefore the most predictive in nature

    MeanDecreseGini: Measure how pure terminal nodes are.
    Again the plot tests results after removing each variable, for decrease in Gini value.
    Variable with the highest value has the highest predictive power

<h2>Performance Evaluation</h2>
<br/>
<h3>Confusionmatrix</h3>
<br/>
```{r ,tidy=TRUE}
#' *Confusionmatrix inbuilt into randomForest()*
    fit[5]
    fit$confusion
#' *Confusionmatrix from caret*
    set.seed(121)
    library(caret)
    confusionMatrix(as.factor(fit$predicted), as.factor(train$Survived))
```

<br/>
<h2>Tuning the nmodel</h2>
<br/>
Creating a new data.frame to contain just the predictors necessary and not all the columns in the original training dataset.
```{r}
    train1 <- data.frame(Pclass = train$Pclass
                         , Survived =train$Survived
                         , Sex=train$Sex,Age=train$Age
                         , SibSp=train$SibSp,Parch=train$Parch
                         , Fare=train$Fare
                         , Embarked=train$Embarked
                         , Title=train$Title
                         , FamilySize=train$FamilySize
                         , FamilyID2=train$FamilyID2)
```
<br/>
Tuning to get the best value for mtry
```{r}
    set.seed(121)
    tunefit=train(as.factor(Survived)~ .
                  , data=train1
                  , method="rf"
                  , metric="Accuracy"
                  , tuneGrid=data.frame(mtry=c(2,3,4)))
    tunefit
    
#' *Prediction*
    prediction=predict(tunefit, newdata=test)
```
</body>