---
title: "Building Interactive Apps with Shiny and R"
author: "Anton"
output:
  pdf_document:
    toc: yes
    toc_depth: '5'
  html_document:
    toc: yes
    theme: united
    toc_float: yes
    toc_depth: 5
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
```

<br/> <br/>

## Introduction

<br/>

::: {.exercise}
We are going to be using the library shiny
:::

```{r , tidy=TRUE}
library(shiny)
```

<br/> Every shiny app has different files:

-   ui.R
-   server.R
-   global.R

### ui.R

Here the User Interface is defined. We define things like inputs and
outputs. Anything we want to show the user on the screen is defined
here.

### server.R

Here we create the objects that are then going to be shown from the ui
file. Here we have functions of inputs and outputs. Any output we create
here gets displayed via the ui file.

### global.R

A shiny app can also have an optional script, called global.R where you
can load static data, define variables and functions and load libraries.
Here you put things that you need to load once in the beginning but do
not need for every step, like e.g. big data loads.

## Shiny inputs

The following code belongs into the ui.R script

Inputs are the way a user can feed data to the shiny app. This can be
text or entering a number, choose date... Below you can see a selection
of possible inputs

```{r , tidy=TRUE}
fluidPage(
  textInput(inputId = "name", label = "Enter your name:", placeholder = "John Smith"),
  numericInput(inputId = "age", label = "Enter your age:", value = 30),
  dateInput(inputId = "birthday", label = "Select your birthday:"),
  selectInput(inputId = "gender", label = "Choose your gender:", choices = c("Male", "Female")),
  sliderInput(inputId = "favNumber", label = "Select your favorite number", min = 0, max = 100, step = 1, value = 50)
)
```

### Text input

For this we use the function textInput

```{r , tidy=TRUE}
fluidPage(

  textInput(inputId = "text" #what this input is going to be called in the whole of the app
            , label = "Enter text" #what is above the box when the user sees the textbox on shiny app, so its basically an introduction
            , value = "" #default value you want in the box, displayed in the bux. The user has to delete the value before entering his own content.
            , placeholder = "Enter text here" #also default value, but user does not have to delete before entering own content
  )
)
```

### Numeric input

For this we use the function numericInput. This is very similar to the
textInput, only that now the user cannot enter a string but only a
number

```{r , tidy=TRUE}
fluidPage(
  
  numericInput(inputId = "numeric"
               , label = "Enter number"
               , value = 5
               , min = 1, max = 10 #defining a range of possible values that can be put in
               , step = 1 #defining the steps in which the range goes up or down
               )
  
)
```

### Checklbox input

For this we use the function checkboxInput(). Here the user can simply
check or uncheck. Additionally we can use the checkboxGroupInput where
one can have multiple choices of boxes.

```{r , tidy=TRUE}
fluidPage(
  
  checkboxInput(inputId = "ceckbox"
                , label = "Klick checkbox"
                , value = TRUE #T or F for whether the box is checked (T) or unchecked (F) as a default
                )
  
  ,checkboxGroupInput(inputId = "checkbiox.multi"
                     , label = "Check or uncheck" #here the label acts for all boxes
                     , choices = c("Choice 1", "Choice 2", "Choice 3") #as a text next to the different boxes
                     )
  
)
```

### Select input

For this we use the function selectInput(). This gives the user the
option of a dropdown menu.

```{r , tidy=TRUE}
fluidPage(
  
  selectInput(inputId = "select"
              , label = "Select your choice"
              , choices = c("Choice 1", "Choice 2", "Choice 3")
              , multiple = F #lets user select multiple if its T and one if its F
              )
  
)
```

### Slider input

For this we use the function sliderInput(). This gives the user the
option of a number line wirh a min and a max.

```{r , tidy=TRUE}
fluidPage(
  
  sliderInput(inputId = "slider"
              , label = "Slide here"
              , min = 1, max = 10
              , value = 5 #default value the slider is set at
              , step = 1 #the value by which each step adjusts by
              )
  
)
```

### Radio button

For this we use the function radioButtons(). This is like a checkbox but
instead of checking yes or no, the user can select one of the choices
displayed on the app.

```{r , tidy=TRUE}
fluidPage(
  
  radioButtons(inputId = "radio"
               , label = "Check your choice"
               , choices = c("Choice 1", "Choice 2", "Choice 3")
               )
  
)
```

### Date Input

For this we use the function dateInputs(). Here the user can select a
date.

```{r , tidy=TRUE}
fluidPage(
  
  dateInput(inputId = "date"
            , label = "Select date"
            , value = "2019-01-01"
              )
  , dateRangeInput(inputId = "dateRange"
                   , label = "Select a range"
                   , start = "2019-01-01"
                   , end = "2019-01-31")
  
)
```

## Shiny Outputs

Outputs are the easiest way to display information to users.

This is defined in the server.R file.

### Text output

Here we use the function renderText() and dave that into output\$text1

```{r , tidy=TRUE}
#' *in our server.R*
function(input, output){
  
  output$text1 <- renderText("This is sample text.")
  output$text2 <- renderText(paste("This is a text", "with a line break", sep ="\n"))
  
}

#' *in our ui.R*
fluidPage(
  
  verbatimTextOutput("text1")
  , verbatimTextOutput("text2")
  
)
```

### Table output

Displaying a table of data in the app

```{r , tidy=TRUE}
#' *in our server.R*
function(input, output){
  
  data("mtcars") #to load data
  
  output$table1 <- renderTable(mtcars) #save table into output$table1
  
}

#' *in our ui.R*
fluidPage(
  
  tableOutput("table1")
  
)
```

### Plot output

Displaying a plot of data in the app

```{r , tidy=TRUE}
#' *in our server.R*
library(ggplot2)

function(input, output){
  
  data("mtcars")
  
  output$plot1 <- renderPlot({
    ggplot(data = mtcars, aes(x=mpg, y=hp)) + geom_point(color = mtcars$cyl)
  }
    
  )
  
}

#' *in our ui.R*
fluidPage(
  
  plotOutput("plot1")
  
)
```

### Data table output

Displaying a table in a more selected way than before in the app.

```{r , tidy=TRUE}
#' *in our server.R*
library(DT)

function(input, output){
  
  data("mtcars")
  
  output$dt1 <- renderDT({
    datatable(mtcars #dataset we are using
              , options = list(pageLength = 5 #how many rows to show default
                                     , lengthMenu = c(5,10,15,20) #letting user decide how many rows he wants to see
                                     )
              )
  }
    
  )
  
}

#' *in our ui.R*
fluidPage(
  
  plotOutput("plot1")
  
)
```

<br/>

## Reactivity

Reactivity is the most important aspect of shiny. It lets the app react
to certain inputs from the users.

### Reactive outputs

::: {.comment}
Ractive outputs allow the output to change based on some input the user
gives you.

A shiny app can be consisting of a UI and a server file, but as well as
a global file in which the libraries are loaded and similar.
:::

```{r , tidy=TRUE}
library(shiny)
library(ggplot2)
data("mtcars")

colChoices <- colnames(mtcars)
```

<br/>

::: {.comment}
:::

```{r , tidy=TRUE}
```

<br/>

::: {.comment}
:::

```{r , tidy=TRUE}
```

<br/>

::: {.comment}
:::

```{r , tidy=TRUE}
```

<br/>

::: {.comment}
:::

```{r , tidy=TRUE}
```

<br/>

::: {.comment}
:::

```{r , tidy=TRUE}
```

<br/>

::: {.comment}
:::

```{r , tidy=TRUE}
```
